---
title: "what i get from the reference 2"
output: html_document
date: "2025-08-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## how to calculate the dispersion of different variables



```{r}
```

In the article, the professor uses the estate price of Bay Area Blues to 
illustrate the effect of the housing crisis.

But in my opinion, i think it is of great importance for us to get method to 
investigate the diversion of the price.

He use the price of different deciles of real estate in Bay area, like 10%, 20%,
30%, ...,90%. use line plot to see its gaps 

And he found that the biggest absolute decline in actual prices occurred at the
expensive end.

However, it is more appropriate to look at the relative prices: how have the 
price change proportionately? 

To do this, the author indexed each decile, dividing the each series by its 
initial price, just as we did for CPI. 

```{r}
#insert the code and show the chart
```

Then the author get the conclusion that the cheapest estates drop the most in 
the proportion.


Another way to illustrate the dispersion is to represent a proportion of the
median house price. 1.2 represents the 20% higher than median. 90% represents 
10% lower than median.

```{r}
#insert the code and show the chart
```



## The website helping me categorize different types of variables 

https://www.data-to-viz.com/